---
editor_options: 
  markdown: 
    wrap: sentence
---

```{r}
n_all_samples = uniqueN(complete_data$gr_sample_id)
earliest_year = year(min(complete_data$date, na.rm = T))
last_year = year(max(complete_data$date, na.rm = T))
samples_distance = uniqueN(distannce_removed$gr_sample_id)
ratio_distance = round(samples_distance/n_all_samples * 100, 0)
n_sites_impact = uniqueN(impact$gr_sample_id)
ratio_impact = round(n_sites_impact/n_all_samples * 100, 0)
time_data = impact[is.na(year(date)) | year(date) >= 2000]
n_sites_time = uniqueN(time_data$gr_sample_id)
ratio_time = round(n_sites_time/n_all_samples * 100, 0)
data_genus_n_genus = ncol(genus_data) - 2

data_all_n_spe = ncol(all_data$spe) - 2
data_all_n_gen = ncol(all_data$gen) - 2
data_all_n_fol = ncol(all_data$foh) - 2
```

# Methods

## Preparation of macroinvertebrate data

<!-- How did we get the data?     -->

<!-- Data curation -->

We compiled a data set of macroinvertebrate samples from rivers throughout Europe.
The data included `r n_all_samples` samples (Figure \@ref(fig:fig-map)) taken between `r earliest_year` and `r last_year` (see Table \@ref(tab:tab-impact-criteria)).
The samples were assigned the BRT-type of the nearest river segment or removed if this segment was further away than 500 m.
As geospatial representation of the BRT we used the data provided by @globevnik2019btfr.
Removing samples with a distance larger than 500 m to the next segment reduced the number of samples to `r samples_distance` (`r ratio_distance`% of all samples).
Next, we removed all samples from sites that were considerably impaired.
To identify these non-reference sites, we used different criteria.
The respective criterion depended on the data set, as some provided detailed information on environmental conditions while other solely included biological information.
For data sets where environmental information was lacking for most sites, we identified impaired sites using the catchment land cover.
Sampling sites from catchments where the combined area of urban and agricultural land exceeded 20% of the catchment area were considered as impaired [@Malaj2014;@Schreiner2016; @Szocs2017].
We used the Catchment Characterization and Modeling database for geospatial data on catchments [@Vogt2007] and the Corine Landcover data from 2018 for landcover information [@clc18]. We provide detailed information on variables and thresholds used to determine the reference sites in table \@ref(tab:tab-impact-criteria). We removed these potentially imparied sites because differences between river types are more pronounced between least-imparied sites [@Verdonschot2006] and because we wanted to compare the typologies capability to delineate reference assemblages with low spatial variability.
Removing impaired sites further decreased the number of samples to `r n_sites_impact` (`r ratio_impact`% of all samples).
Lastly, we only considered the `r n_sites_time` samples that were taken in or after the year 2000 (`r ratio_time`% of all samples).

```{r fig-map, fig.cap = "Map of sampling sites. The first map (too far) shows sites that were more than 500 meters removed from the closesed river segement. The second map (impacted) shows sites that were categorized as not in referene condition. The third map (before 2000) shows sites that were sampled before 2000. The fourth map (final) shows all samples that were used for the analyses."}

img <- readPNG("figures/map_for_paper.png")
grid.raster(img)
```

<!-- How did we harmonize data sets.  -->

The data originated from different sources and required adjustments to ensure taxonomic consistency.
<!-- We used the Global Biodiversity Information Facility (GBIF) to check whether any of our species were considered synonyms and if so changed their names to the accepted names suggested in GBIF. --> We used the Global Biodiversity Information Facility (GBIF) to replace synonyms with accepted names and to assign a taxonomic level to each observation.
To integrate the different data types we transformed all data to presence-absence.
Next, we restricted observations to the following classes: Clitellata, Insecta, Malacostraca, Bivalvia or Gastropoda.
<!-- We used different approaches for the different aims of this study. --> 
Hereafter, we differentiate between two data sets: *data genus* and *data all*.
We used *data genus* to compare the different classifications and to derive indicator taxa.
This data set was harmonized at the genus level.
All observations with a lower taxonomic resolution (i.e. family, order, etc.) were removed.
This resulted in a data set with `r data_genus_n_genus` genera.
We used *data all* to derive the typical assemblages.
This data set has a mixed taxonomic levels.
Taxa can be represented at species, genus, family, order or sublcass level. 
<!-- Instead, we aimed to represent each taxon with a taxonomic level that would be optimal for it, given the data available for the river type.  --> <!-- This increased the number of observations that we could use but decreased the comparability between river types.  --> <!-- Since river types do not influence the typical assemblages of other river types, comparability was not necessary.  --> <!-- To this end, consistency between river types was not as important because other river types do not exert large influence on the selection process.  --> <!-- To derive the typical assemblages  --> <!-- Hence we decided to use varying taxonomic levels instead of a fixed one. --> <!-- This way some taxa might be represented at family level and others at species level, depending on the resolution of the data. --> <!-- *Oligochaetes*, for example, are usually only determined to subclass level. --> <!-- In a setting with one common taxonomic level (e.g. genus) they would have to be omitted if this level would be higher than subclass. --> <!-- By using taxon- and river type specific levels that take low-resolved taxa into account, we can thus use more of the data. --> <!-- Using different levels poses the challenge of finding an optimal level for each taxon given a data set and river type. --> <!-- We established the optimal level with a hierarchical approach. --> <!-- In the following, a higher taxonomic level refers to levels with higher resolution, i.e. species is the highest taxonomic level and kingdom the lowest. --> <!-- For each taxon, we calculated the percentage of observations represented at each higher level. --> <!-- For example, 4.12% of observations from the order *Lepidoptera* are at the species level, 74.77% at the genus level, 7.75% at the family level, and 13.35% at the order level. --> <!-- Now given a threshold X, we hold a taxon to be optimally represented at a certain taxonomic level if less than X% are represented by higher levels. --> <!-- For example, *Lepidoptera* would be represented on order level if X \> 4.12% + 74.77% + 7.75% = 86.64%. --> <!-- We used 50% as a threshold. --> <!-- Observations that had a lower resolution than what was proposed as optimal, were included at their respective level.  --> 
We aimed to represent each taxon with a taxonomic level that would be optimal for it, given the data available for the river type.
We aimed to represent each taxon at a taxonomic level that would be good trade-off between taxonomic resolution and comparability between data sets. 
We thereby increased the number of observations that we could use but decreased the comparability between river types.
Since river types do not influence the typical assemblages of other river types, comparability was not necessary.
Using different levels posed the challenge of finding an optimal level for each taxon given a data set and river type.
We established the optimal level with a hierarchical approach.
For each taxon, we calculated the percentage of observations represented at each higher taxonomic level.
If this value was below 50 %, we harmonized all observations to this, optimal, level.
If it was above 50%, we repeated the procedure with the next higher level.
We also included all observations with lower than optimal resolutions.
The final *data all* contained `r data_all_n_spe` taxa represented at the species level, `r data_all_n_gen` taxa represented at genus level and `r data_all_n_fol` taxa represented at family or lower taxonomic levels.

<!-- At which types did we look?  -->

After removing samples that did not meet our criteria, we did not have sufficient data to adequately represent all river types.
We considered the data for a river type insufficient if it contained too few sites (\< 20) or the distribution of sites was not representative of the river type.\
Maps of the sampling locations for the individual river types are provided in the Supplementary Materials.
We were not able to represent the river types: BRT6, 7, 12, 13, 17, 19 and 20.
Samples from these river types were removed.
<!-- The river types we were able to represent are: RT1, RT2, RT3, RT4, RT8, RT9, RT10, RT11, RT14, RT15, RT16 and RT18. -->

## Comparison of typoglogies

To evaluate the two typologies proposed by @LycheSolheim2019, we compared them with the three other typologies: i) the k-means typology of the global river classification framework [GloRiC, @OuelletDallaire2019]; ii) the freshwater ecoregions proposed by @illies1978limnofauna; and iii) the biogeographical regions proposed by the European Environmental Agency [BGR, @EuropeanEnvironmentalAgencyEEA2016].
These typologies represent two different kinds of freshwater typologies: reach-based typologies (BRT and GloRiC) and regional typologies (Illies and BGR).
Reach-based typologies classify single river reaches.
The instances of individual types are not spatially contiguous but can be far apart and are commonly very close to instances of different types.
Regional typologies classify large contiguous areas and there always is only one instance of each type.
The different types are only adjacent at the regional margins.
In addition to these five established typologies, we also evaluated BRTred, a reduced version of BRT20.
In BRTred, we combined the river types with redundant typical assemblages (see section Indicator and typical assemblages for more details).
Only genus-level data (*data genus*) were used for this analysis.
We removed all observations from types that were represented by less than twenty sites from the data set.
See table \@ref(tab:tab-types) for the remaining types per typology.

In order to be able to judge the results, we created two additional typologies.
As an upper bound of what to expect, we created a classification of biological data using flexible beta clustering [@lance1967general] with the $\beta$ parameter set to -0.25 and the common parameterization $(\alpha_1 = \alpha_2 = \alpha, \beta = (1-\alpha)/2, \gamma = 0)$.
We chose the number of groups that maximized the average silhouette width (9).
Since this typology is not constrained by environment or space but only represents patterns in the biological data, we expect it to delineate more clearly between different biological assemblages than the other typologies.\
As a lower bound, we created 100 random partitions of the data.
For each partition, we first drew the number of classes as a random variable from the interval between the lowest number of types in any of the typologies tested (6 in BGR) and the highest number (14 in BRT20 and GloRiC).
Then we assigned each observation randomly to one of the groups.

We calculated four cluster quality metrics for each typology: the average silhouette width, the Calinski-Harabasz index, an indicator value score and the classification strength.
We used the Jaccard distance between sites as a distance metric.

<!-- What is ASW  -->

The average silhouette width [ASW, @Kaufmann1990] is computed as $$ASW = \frac{1}{n}\Sigma_{i=1}^n \dfrac{a_i-b_i}{max(a_i,b_i)}$$ where $a_i$ is the average dissimilarity of point $i$ to points from its type, $b_i$ is the average dissimilarity of point $i$ to points from the closest other type and $n$ is the number of observations.
Positive values indicate that on average points are more similar to observations from their own type than to those of the most similar one.
Therefore, high scores imply better typologies.
@Lengyel2019 recently proposed a generalized version of the ASW.
By using the arithmetic average to compute $a_i$ and $b_i$, spherical clusters are assumed to be optimal.
Using a generalized mean instead, we can flexibly adjust our validity metric to put a stronger emphasis on compactness ($a_i$) or separation ($b_i$).
The generalized mean of degree $p$ ($M^p$) is computed as: $$ M^p (\mathbf{x}) = \bigg( \dfrac{1}{n} \Sigma_{i=1}^n x_i^p\bigg)^{1/p} $$ The generalized mean can take the value of common summary statistics such as the minimum ($p=-\infty$), the maximum ($p=\infty$) or the harmonic mean ($p=-1$).
For example, for $p=-\infty$ the silhouette width is the difference between the minimum distance of observation $i$ to any other observation from the same type and the minimum distance from that observation to any observation from the next closest type.
This perspective excludes outliers and values separation over compactness.
The weighting shifts towards compactness as we increase $p$.
We evaluated the silhouette width for $p\ \in \{-\infty,-2,-1,1,2,\infty\}$.
If not further specified, ASW refers to the common average silhouette width (i.e. $p = 1$) in the remainder of the text.
<!-- use cases -->

<!-- What is Calinski-Harabasz? -->

The Calinksi-Harabasz Index [CH, @calinski1974dendrite] is computed as $$CH = \dfrac{BGSS}{WGSS} \times \dfrac{n-k}{k-1}$$ where $BGSS$ is the squared sum of distances between group centroids and the overall centroid (between group sum-of-squares), $WGSS$ is sum of squares of distances between observations of one group (within group sum-of-squares), $k$ is the number of clusters. High values indicate that variation within types is smaller than between types As the second term controls for the degrees of freedom, it can be understood as an analog to the F-Statistic. The algorithm assumes Euclidean data but good performance with a similar metrics was shown for binary data in the context of fMRI-scans [@Dimitriadou2004].
<!-- use cases -->

<!-- What is the indicator value score? -->

The indicator value score (IVS) is based on the indicator value (IndVal) proposed by @dufrene1997species.
The IndVal itself will be explained in detail below.
Here, we only note that we used 999 permutations to compute *p*-values and in contrast to the latter application did not control the family-wise error rate.
IVS is the fraction of taxa that are statistically significant indicators (at a significance level of 0.01) for some type of a typology.
Higher scores indicate a better classification.

<!-- What is ISAMIC -->

```{=html}
<!--The Indicator Species Analysis Minimizing Intermediate Constancies [ISAMIC, @Roberts2020] measures the constancy of a taxon within clusters.
A taxon has a high constancy if it occurs in all or no samples from a cluster.
The constancy is minimized if it occurs in half of the samples.
The ISAMIC is computed as: 
$$ISAMIC = \dfrac{\Sigma_{i=1}^M 2 \times(\Sigma_{k=1}^K |c_{ik} - 0.5|/K)}{M}$$
, where $M$ is the number of taxa and $c_{ik}$ the constancy of taxon $i$ in cluster $k$.
The statistics has been used to evaluate typologies before [e.g. @Aho2008; @Lotter2013; @Roberts2015].
Again, higher values indicate sharper classifications. 
For ISAMIC and IVS, we assume especially high values for the biological typology, as the same data that are tested, were also used to derive the classification [@Caceres2009].-->
```
<!-- What is Classification strength -->

Lastly, we computed the classification strength [CS, @VanSickle1997]. Classification strength is the difference between mean within cluster similarity ($\overline{W}$) and mean between cluster ($\overline{B}$) similarity. As such it ranges between 0 ($\overline{W} = \overline{B}$) and 1 ($\overline{B} = 0$), where higher values indicate a stronger classification. A similar and recently applied metric is the partition analysis [@Roberts2019] which is the ratio of $\overline{W}$ and $\overline{B}$. Here, we opted for the former as is has been used in similar analyses [e.g. @Gerritsen2000; @VanSickle2000; @Vasconcelos2013] which enables us to directly compare our results with those of previous studies.
Note that results did not differ qualitatively between CS and partition analysis (data not shown).
Similarities were calculated as 1 - Jaccard distance.

<!--What is within cluster sum of squares?  -->

<!-- The within cluster sum-of-squares (wcss) is computed as  -->

<!-- $$wcss = \Sigma_{k=1}^{K} \frac{\Sigma_{i=1}^{n_k}\Sigma_{j=1}^{n_k} \frac{d_{mij^2}}{n_k}}{2}$$ -->

<!-- , where $d_{ij}$ is the dissimilarity of observations $i$ and $j$ and $n_k$ is the number of observation in cluster $k$. -->

<!-- If $d$ is a Euclidean distance this is equal to the objective function of k-means clustering. -->

Based on these four cluster criteria, each typology was assigned a score.
We used these scores to evaluate the overall performance of the typologies.
The typology that performed best in some metric received 6 points, the second 5, the third 4 and so on.
Differences smaller than 5% of the range between biological and random partitions were regarded as ties.
When typologies were tied, they all received the mean of their untied scores.
For example, if three typologies are tied for the first place, instead of 6, 5 and 4 points respectively, each receives a score of 5.

## Indicator and typical assemblages

<!-- How did we derive indicator species? -->

Both indicator and typical assemblages were derived for BRT20.
We used the IndVal approach of @dufrene1997species to identify indicator taxa. For this analysis, we used *data genus* which consists of genus-level presence-absence data. The IndVal can be understood as the product of the two quantities $A$ and $B$. For our purposes, $A$ is the relative number of observations of taxon $i$ that are within type $j$. It was initially described as specificity [@dufrene1997species] but is better understood as concentration [@Podani2010] because it is independent of the total number of types. $B$ is the relative frequency with which species $i$ occurs in type $j$. The maximum score is assigned to a species that only occurs in one type ($A=1$) and occurs in all samples of that type ($B=1$). Here, we used the group-equalized version of the IndVal which accounts for the fact that the number of samples differs between types. The statistical significance of the IndVal statistic was assessed with a permutation test that computes IndVal values for random permutations of sites and types and compares the observed IndVal against this empirical distribution. We used $2*10^5$ permutations. We controlled the family-wise error rate with Bonferroni's correction [@bonferroni1935calcolo]. This correction is conservative and has low statistical power [@Nakagawa2004; @Garcia2004].
Hence, we regard the hypothesis, that a genus is an indicator for a broad river type, strongly supported by our data if we find the indication to be statistically significant.
We used 0.05 as the base significance level.

<!-- How did we derive typical assemblages?  -->

These indicator species provide valuable insight into the communities but miss the ubiquitous generalist species that occur in many types (tramp species *sensu* @McGeoch2002).
Even if these taxa are common within a type (high $B$), they will typically have low concentrations in most types (low $A$) and hence low and statistically non-significant indicator values.
Hence, the indicator assemblages do not represent a typical sample, in the sense that these taxa can reasonably be expected to occur in samples of the type.
We derived such typical assemblages by setting explicit thresholds for $B$.
We used *data all* to derive typical assemblages.
These data have different taxonomic levels and we set different thresholds for different taxonomic levels.
All species that occurred in more than 25% of samples of a river type (i.e. $B > 0.25$) were considered typical.
The value increased to 33% for genera and 66% for families or lower taxonomic levels.
<!--We also included species with very high concentrations ($A > 0.9$) but only if they also occurred in more than 5% of samples, to prevent singletons and rare species from exerting undue influence. --> Others used alterations of the indicator value before, even if this was implicit.
For example, @wagner2001quantifying considered only the A component.
@McGeoch2002 discusses the advantage of different combinations of A and B.

After deriving typical assemblages, we evaluated their similarity using the Jaccard similarity.
A similarity of 0.5 indicates that half of the taxa in the combined taxa pool occur in both typical assemblages.
If the similarity between two typical assemblages exceeded 0.8, we deemed the river types redundant and combined them.
For example, the broad river types BRT2 and BRT3 (medium to large and very small to small siliceous lowland rivers) might be found to be redundant and combined into BRT2_3 (very small to large siliceous lowland rivers).
All sites belonging to either of these river types would also be reclassified and the typical assemblages would be derived again.
This is repeated until no similarity exceeded 0.8.
We did not do the same with indicator assemblages as they are explicitly being optimized for being different from one another (through the equal weighting of A and B).
This way we can evaluate whether the ad-hoc combinations of river types used to derive the BRT12 typology are justified by biological homogeneity.

<!-- We evaluated the average taxonomic resolution of the data available for each river type. -->

<!-- To this end, we assigned a score of three to each observed species, two to each genus and one to each family or lower resolved taxon and computed the average score for each river type. -->

<!-- How did we assign traits?  -->

<!-- We used a harmonized trait data-base of European macroinvertebrates (Kunz pers comm) to assign traits values to the indicator and typical assemblages.  -->

<!-- The traits are fuzzy-coded [@chevene1994fuzzy], which allow single species to express tendencies towards different trait modalities.  -->

<!-- Instead of assigning a single trait modality, fuzzy-coding encodes affinities for each modality, in our case ranging from 0 (low affinity) to 1 (high affinity). -->

<!-- This can represent metaphoetetic taxa or genera with variation among species.  -->

<!-- We considered seven traits: feeding mode, respiration strategy, voltinism, locomotion, oviposition, body size and body form.  -->

<!-- If the traits for a taxon were not available in the data-base we used that of a lower taxonomic level, if available.  -->

<!-- Taxa that could not be assigned traits were removed from the trait analysis.  -->

<!-- We computed mean modality affinities for each trait for typical and indicator assemblages.  -->

<!-- Software  -->

## Software

All computations were conducted in the R Statistical Environment v. 4.0.3 [@RCT2020]. Data were prepared using data.table 1.14.0 [@Dowle2021], tidyverse packages [@Wickham2019] and taxize 0.9.98 [@ScottChamberlain2013; @Chamberlain2020]. Geospatial analyses were conducted using sf [@Pebesma2018]. Clusters were created and evaluated with fpc [@Hennig2020], indicspecies [@Caceres2009], labdsv [@Roberts2019], optpart [@Roberts2020]. Generalized silhouette widths were computed with the R functions provided in the supplementary materials of @Lengyel2019. Figures and maps were created with ggplot2 [@Wickham2016] and tmap [@Tennekes2018].
