---
editor_options: 
  markdown: 
    wrap: sentence
---

```{r}
n_all_samples = uniqueN(complete_data$gr_sample_id)
earliest_year = year(min(complete_data$date, na.rm = T))
last_year = year(max(complete_data$date, na.rm = T))
samples_distance = uniqueN(distannce_removed$gr_sample_id)
ratio_distance = round(samples_distance/n_all_samples * 100, 0)
n_sites_impact = uniqueN(impact$gr_sample_id)
ratio_impact = round(n_sites_impact/n_all_samples * 100, 0)
time_data = impact[is.na(year(date)) | year(date) >= 2000]
n_sites_time = uniqueN(time_data$gr_sample_id)
ratio_time = round(n_sites_time/n_all_samples * 100, 0)
data_genus_n_genus = ncol(genus_data) - 2

data_all_n_spe = ncol(all_data$spe) - 2
data_all_n_gen = ncol(all_data$gen) - 2
data_all_n_fol = ncol(all_data$foh) - 2
```

# Methods

## Preparation of macroinvertebrate data

<!-- How did we get the data?     -->

<!-- Data curation -->

We compiled a data set of macroinvertebrate samples from rivers throughout Europe.
The data included `r n_all_samples` samples (Figure \@ref(fig:fig-map)) taken between `r earliest_year` and `r last_year` (see Table \@ref(tab:tab-impact-criteria)).
The samples were assigned the BRT-type of the nearest river segment or removed if this segment was further away than 500 m.
As geospatial representation of the BRT we used the data provided by @globevnik2019btfr.
Removing samples with a distance larger than 500 m to the next segment reduced the number of samples to `r samples_distance` (`r ratio_distance`% of all samples).
Next, we removed all samples from sites that were considerably impaired.
To identify these non-reference sites, we used different criteria.
The respective criterion depended on the data set, as some provided detailed information on environmental conditions while other solely included biological information.
For data sets where environmental information was lacking for most sites, we identified impaired sites using the catchment land cover.
Sampling sites from catchments where the combined area of urban and agricultural land exceeded 20% of the catchment area were considered as impaired [@Malaj2014; @Schreiner2016; @Szocs2017]. We used the Catchment Characterization and Modeling database for geospatial data on catchments [@Vogt2007] and the Corine Landcover data from 2018 for landcover information [@clc18]. We provide detailed information on variables and thresholds used to determine the reference sites in table \@ref(tab:tab-impact-criteria). We removed these potentially impaired sites because differences between river types are more pronounced between least-impaired sites [@Verdonschot2006] and because we wanted to compare the typologies capability to delineate reference assemblages with low spatial variability.
Removing impaired sites further decreased the number of samples to `r n_sites_impact` (`r ratio_impact`% of all samples).
Lastly, we only considered the `r n_sites_time` samples that were taken in or after the year 2000 (`r ratio_time`% of all samples).

The resulting data set had sufficient data to represent 13 of 20 broad river types.
We considered our samples to be representative for their river type, if we had more than 20 sites, which were distributed evenly across the instances of the river type.\
Maps of the sampling locations for the individual river types are provided in the Supplementary Materials.
We had insufficient samples of the river types: BRT6, 7, 12, 13, 17, 19 and 20, consequently, these were omitted from the further analysis.

```{r fig-map, fig.cap = "Map of sampling sites. The first map (too far) shows sites that were more than 500 meters removed from the closesed river segement. The second map (impacted) shows sites that were categorized as not in referene condition. The third map (before 2000) shows sites that were sampled before 2000. The fourth map (final) shows all samples that were used for the analyses."}

img <- readPNG("figures/map_for_paper.png")
grid.raster(img)
```

<!-- How did we harmonize data sets.  -->

The data originated from different sources and required adjustments to ensure taxonomic consistency.
We used the Global Biodiversity Information Facility (GBIF) to replace synonyms with accepted names and to assign a taxonomic level to each observation.
To integrate the different data types we transformed all data to presence-absence.
Next, we restricted observations to the following classes: Clitellata, Insecta, Malacostraca, Bivalvia and Gastropoda.
When harmonizing taxonomies across data sets, there is a trade-off between comparability among data sets and taxonomic resolution.
If we employ a single taxonomic level (e. g. genus) for all taxa across river types, we have to agglomerate observations with finer resolutions and omit those with coarser resolutions.
The resulting data set would be appropriate for comparisons between river types because taxa have consistent representations across river types.
If we vary the taxonomic resolution between river types and taxa, we artificially increase the dissimilarity between river types but attain the best available representation for the individual river types.
To compare freshwater typologies, we used the former approach and harmonized all observations at the genus level, omitting all observations with coarser taxonomic levels.
This resulted in a data set with `r data_genus_n_genus` genera.
We established typical assemblages per river type independently, which allowed us to set an appropriate taxonomic level per river type and taxon.
This was done using a hierarchical approach.
For each taxon, we calculated the percentage of observations represented at each finer taxonomic resolution.
If this value was below 50 %, we harmonized all observations to this level.
If it was above 50%, we repeated the procedure with the next finer level.
We also included all observations with coarser than optimal resolutions.
The resulting data set contained `r data_all_n_spe` taxa represented at the species level, `r data_all_n_gen` taxa represented at genus level and `r data_all_n_fol` taxa represented at family or lower taxonomic levels.

## Comparison of typologies

@LycheSolheim2019 proposed two versions of the BRT typology.
BRT20 differentiates between 20 different river types, some of which are very rare.
To reduce the number of broad river types @LycheSolheim2019 merged types which very few instances with types that covered many water bodies and match the rare type in at least two type descriptors.
The resulting typology (BRT12) has twelve types.
As previous publications have used both (e.g. \@Posthuma2020 used BRT12 and \@Poikane2019 used BRT20) we consider both.

To evaluate the two versions of BRT, we compared them with the three other typologies: i) the k-means typology of the global river classification framework [GloRiC, @OuelletDallaire2019]; ii) the freshwater ecoregions proposed by @illies1978limnofauna; and iii) the biogeographical regions proposed by the European Environmental Agency [BGR, @EuropeanEnvironmentalAgencyEEA2016].
These typologies can be categorized into: reach-based typologies (BRT and GloRiC) and regional typologies (Illies and BGR).
Reach-based typologies classify single river reaches.
The instances of individual types are not spatially contiguous and can be far apart.
They are commonly very close to instances of different types.
By contrast, regional typologies classify large contiguous areas and there is only one instance per type.
Different types are only adjacent at the boundaries between two types.
As with BRT before, we retained those types for which we had more than 20 samples.
See table \@ref(tab:tab-types) for the remaining types in each typology.

For each typology, we computed four cluster validity metrics, each emphasizing a different aspect of cluster quality: the average silhouette width, the Calinski-Harabasz index, an indicator value score and the classification strength.
We used the Jaccard distance between sites as a distance metric.

As a reference frame for these for metrics, we created two additional typologies.
To have an upper bound of what to expect, we clustered the sampling sites with an agglomerative clustering method.
This typology is constrained neither by environment nor by space but only by the biological data.
Thus, we expect it to delineate more clearly between different biological assemblages than the other typologies and to achieve better cluster validity scores.
We will refer to this typology as the biological typology. 
All agglomerative clustering methods begin with as many clusters as observations and then they iteratively combine the most similar observations into clusters until only one cluster remains.
The approach differ in how this similarity is computed.
@lance1967general clarified that all approaches are just different parametrizations of one overarching method: flexible beta clustering.
Flexible beta clustering has four parameters but it is common to only specify one, $\beta$, while the other three parameters are derived from it.
We compared different final group sizes and $\beta$ parameters to find the optimal solution and refer the interested reader to the supplementary materials for more detailed information.

As a lower bound of what to expect, we created 100 random typologies.
Differences between types are purely random and therefore should be no systematic differences between the types.
We hence expect these random typologies to receive worse cluster validity scores than all the other tested typologies.
For each random typology, we first drew the number of types as a random variable from the interval between the lowest number of types in any of the typologies (6 in BGR) and the highest number (14 in BRT20 and GloRiC).
Then we assigned each observation randomly to one of the groups.

Here, we provide a short primer on each cluster validity metric we used.
More extended information is provided in the supplementary materials.
<!-- What is ASW  --> The average silhouette width [ASW, @Kaufmann1990] the average of all silhouette widths.
A silhouette width is calculated for each sample.
It is the difference between the mean distance of sample $i$ to all samples from the most similar type and the mean distance of sample $i$ to all other samples from the same type.
The silhouette width is scaled so it lies between -1 and 1.
Values close to 1 indicate that most samples are more similar to samples from their own type than to those of other types.
Values around zero suggest that the sample lies between two types and values close to -1 signal that the sample is more similar to samples from another types than its own.
In addition to the commonly applied ASW, we also computed several modified versions. 
They summarize the distance in $a_i$ and $b_i$ with the generalized mean [@Lengyel2019]. 
The generalized mean has one modifiable parameter, $p$, and for $p=1$ the modifiable mean is equal to the arithmetic mean. 
As we increase $p$ the weight of outlying samples increases.  
<!-- What is Calinski-Harabasz? -->
The Calinski-Harabasz Index [CH, @calinski1974dendrite] is the scaled ratio of the distances between types and the distances within types.
Higher values indicate that distances between types are larger than those within types and thus that samples are more similar to samples from their own type than to samples from other types.  
<!-- What is the indicator value score? -->
The indicator value score (IVS) is based on the indicator value (IndVal) proposed by @dufrene1997species.
The IndVal itself will be explained in detail below.
Here, we only note that we used 999 permutations to compute *p*-values and ignored the family-wise error rate.
IVS is the fraction of taxa that are statistically significant indicators (at a significance level of 0.01) for some type of a typology.
Higher scores indicate a better classification.

<!-- What is Classification strength -->
Lastly, we computed the classification strength [CS, @VanSickle1997]. Classification strength is the difference between mean within cluster similarity ($\overline{W}$) and mean between cluster ($\overline{B}$) similarity. We calculated similarities as 1 - Jaccard distnace. The CS ranges between 0 ($\overline{W} = \overline{B}$) and 1 ($\overline{B} = 0$), where higher values indicate a stronger classification. CS has been used in similar analyses [e.g. @Gerritsen2000; @VanSickle2000; @Vasconcelos2013], which enables us to directly compare our results with those of previous studies.

We ranked the typologies based on each of these four cluster validity metrics,
from 1 to 5 relating to the worst and best performances respectively.
Differences smaller than 5% of the range between the biological and the random typology were regarded as ties.
When typologies were tied, they all received the mean of their untied scores.
For example, if three typologies are tied for the first place, instead of 6, 5 and 4 points, each receives a score of 5.

## Indicator and typical assemblages

<!-- How did we derive indicator species? -->

Both indicator and typical assemblages were derived for BRT20.
We used the IndVal approach of @dufrene1997species to identify indicator taxa. For this analysis, we used *data genus* which consists of genus-level presence-absence data. The IndVal can be understood as the product of the two quantities $A$ and $B$. For our purposes, $A$ is the relative number of observations of taxon $i$ that are within type $j$. It was initially described as specificity [@dufrene1997species] but is better understood as concentration [@Podani2010] because it is independent of the total number of types. $B$ is the relative frequency with which species $i$ occurs in type $j$. The maximum score is assigned to a species that only occurs in one type ($A=1$) and occurs in all samples of that type ($B=1$). Here, we used the group-equalized version of the IndVal which accounts for the fact that the number of samples differs between types. The statistical significance of the IndVal statistic was assessed with a permutation test that computes IndVal values for random permutations of sites and types and compares the observed IndVal against this empirical distribution. We used $2*10^5$ permutations. We controlled the family-wise error rate with Bonferroni's correction [@bonferroni1935calcolo]. This correction is conservative and has low statistical power [@Nakagawa2004; @Garcia2004].
Hence, we regard the hypothesis, that a genus is an indicator for a broad river type, strongly supported by our data if we find the indication to be statistically significant.
We used 0.05 as the base significance level.

<!-- How did we derive typical assemblages?  -->

These indicator species provide valuable insight into the communities but miss the ubiquitous generalist species that occur in many types (tramp species *sensu* @McGeoch2002).
Even if these taxa are common within a type (high $B$), they will typically have low concentrations in most types (low $A$) and hence low and statistically non-significant indicator values.
Hence, the indicator assemblages do not represent a typical sample, in the sense that these taxa can reasonably be expected to occur in samples of the type.
We derived such typical assemblages by setting explicit thresholds for $B$.
We used *data all* to derive typical assemblages.
These data have different taxonomic levels and we set different thresholds for different taxonomic levels.
All species that occurred in more than 25% of samples of a river type (i.e. $B > 0.25$) were considered typical.
The value increased to 33% for genera and 66% for families or lower taxonomic levels.
<!--We also included species with very high concentrations ($A > 0.9$) but only if they also occurred in more than 5% of samples, to prevent singletons and rare species from exerting undue influence. --> Others used alterations of the indicator value before, even if this was implicit.
For example, @wagner2001quantifying considered only the A component.
@McGeoch2002 discusses the advantage of different combinations of A and B.

After deriving typical assemblages, we evaluated their similarity using the Jaccard similarity.
A similarity of 0.5 indicates that half of the taxa in the combined taxa pool occur in both typical assemblages.
If the similarity between two typical assemblages exceeded 0.8, we deemed the river types redundant and combined them.
For example, the broad river types BRT2 and BRT3 (medium to large and very small to small siliceous lowland rivers) might be found to be redundant and combined into BRT2_3 (very small to large siliceous lowland rivers).
All sites belonging to either of these river types would also be reclassified and the typical assemblages would be derived again.
This is repeated until no similarity exceeded 0.8.
We did not do the same with indicator assemblages as they are explicitly being optimized for being different from one another (through the equal weighting of A and B).
This way we can evaluate whether the ad-hoc combinations of river types used to derive the BRT12 typology are justified by biological homogeneity.

<!-- We evaluated the average taxonomic resolution of the data available for each river type. -->

<!-- To this end, we assigned a score of three to each observed species, two to each genus and one to each family or lower resolved taxon and computed the average score for each river type. -->

<!-- How did we assign traits?  -->

<!-- We used a harmonized trait data-base of European macroinvertebrates (Kunz pers comm) to assign traits values to the indicator and typical assemblages.  -->

<!-- The traits are fuzzy-coded [@chevene1994fuzzy], which allow single species to express tendencies towards different trait modalities.  -->

<!-- Instead of assigning a single trait modality, fuzzy-coding encodes affinities for each modality, in our case ranging from 0 (low affinity) to 1 (high affinity). -->

<!-- This can represent metaphoetetic taxa or genera with variation among species.  -->

<!-- We considered seven traits: feeding mode, respiration strategy, voltinism, locomotion, oviposition, body size and body form.  -->

<!-- If the traits for a taxon were not available in the data-base we used that of a lower taxonomic level, if available.  -->

<!-- Taxa that could not be assigned traits were removed from the trait analysis.  -->

<!-- We computed mean modality affinities for each trait for typical and indicator assemblages.  -->

<!-- Software  -->

## Software

All computations were conducted in the R Statistical Environment v. 4.0.3 [@RCT2020]. Data were prepared using data.table 1.14.0 [@Dowle2021], tidyverse packages [@Wickham2019] and taxize 0.9.98 [@ScottChamberlain2013; @Chamberlain2020]. Geospatial analyses were conducted using sf [@Pebesma2018]. Clusters were created and evaluated with fpc [@Hennig2020], indicspecies [@Caceres2009], labdsv [@Roberts2019], optpart [@Roberts2020]. Generalized silhouette widths were computed with the R functions provided in the supplementary materials of @Lengyel2019. Figures and maps were created with ggplot2 [@Wickham2016] and tmap [@Tennekes2018].

```{=html}
<!-- OLD CODE 
Hereafter, we differentiate between two data sets: data genus and data all. We used data genus to compare the different classifications and to derive indicator taxa. This data set was harmonized at the genus level. All observations with a lower taxonomic resolution (i.e. family, order, etc.) were removed. This resulted in a data set with r data_genus_n_genus genera. We used data all to derive the typical assemblages. This data set has a mixed taxonomic levels. Taxa can be represented at species, genus, family, order or sublcass level. We aimed to represent each taxon with a taxonomic level that would be optimal for it, given the data available for the river type. We aimed to represent each taxon at a taxonomic level that would be good trade-off between taxonomic resolution and comparability between data sets. We thereby increased the number of observations that we could use but decreased the comparability between river types. Since river types do not influence the typical assemblages of other river types, comparability was not necessary. Using different levels posed the challenge of finding an optimal level for each taxon given a data set and river type. We established the optimal level with a hierarchical approach.
-->
```
