---
editor_options: 
  markdown: 
    wrap: sentence
---

```{r}
n_all_samples = uniqueN(complete_data$gr_sample_id)
earliest_year = year(min(complete_data$date, na.rm = T))
last_year = year(max(complete_data$date, na.rm = T))
samples_distance = uniqueN(distannce_removed$gr_sample_id)
ratio_distance = round(samples_distance/n_all_samples * 100, 0)
n_sites_impact = uniqueN(impact$gr_sample_id)
ratio_impact = round(n_sites_impact/n_all_samples * 100, 0)
time_data = impact[is.na(year(date)) | year(date) >= 2000]
n_sites_time = uniqueN(time_data$gr_sample_id)
ratio_time = round(n_sites_time/n_all_samples * 100, 0)
data_genus_n_genus = ncol(genus_data) - 2

data_all_n_spe = ncol(all_data$spe$brt12) - 2
data_all_n_gen = ncol(all_data$gen$brt12) - 2
data_all_n_fol = ncol(all_data$foh$brt12) - 2
```

# Methods

## Preparation of macroinvertebrate data

<!-- Data curation -->
We compiled a data set of macroinvertebrate samples from rivers throughout Europe (Table \@ref(tab:tab-impact-criteria)).
A visual summary of the steps described below is provided in Figure \@ref(fig:fig-workflow).
The data included `r n_all_samples` samples taken between `r earliest_year` and `r last_year`.
We assigned the BRT-type of the closest river segment to samples if said segments was within 500 m of the sample. 
If a sample was further than 500 m removed from the next segment, we omitted it. 
Removing samples with a distance larger than 500 m to the next segment reduced the number of samples to `r samples_distance` (`r ratio_distance`% of all samples).
Next, we removed all samples from sites that were considerably impaired.
Impairment was evaluated based on the definition of the data set (see Table 1) and where this information was lacking, based on catchment land cover.
Sampling sites from catchments where the combined area of urban and agricultural land exceeded 20% of the catchment area were considered impaired [@Malaj2014; @Schreiner2016; @Szocs2017;@Wolfram2021].
Though referring to catchment land cover for all sites would result in a more homogeneous processing, we referred to specific information where available to increase precision in the evaluation of impairment.
We removed these potentially impaired sites because differences between river types are more pronounced between least-impaired sites [@Verdonschot2006] and because we wanted to compare the typologies' capability to delineate reference assemblages with low spatial variability.
Removing impaired sites further decreased the number of samples to `r n_sites_impact` (`r ratio_impact`% of all samples).
Lastly, we only considered the `r n_sites_time` samples that were taken in or after the year 2000 (`r ratio_time`% of all samples).
The spatial distribution of the final data set is shown in Figure \@ref(fig:fig-map-used). 
We provide maps of the spatial distribution of removed sites in the supplementary information.  
\newpage

\blandscape

```{r tab-impact-criteria}
impact = matrix(nrow = 14, ncol = 6)
impact %<>% as.data.frame()
names(impact) = c("Data set", "Area", "Time span", "Number of Samples", "Exclusion criterion", "Reference")
impact$`Data set` = c("Monitoring data from German federal agencies", 
                      "Ebro Hydrographic Confederation", 
                      "Naiades",
                      "Cantabria",
                      "Picos de Europa",
                      "Ecosurv",
                      "RIVPACS database",
                      "STAR",
                      "WISER",
                      "AQUEM Sweden",
                      "Koutajoki drainage basin",
                      "Monitoring data from the RCS national network managed by the French water agencies",
                      "Monitoring data from Dutch regional water authorities ",
                      "Monitoring data from the German Federal state of Hesse")

impact$`Time span` = c("1968 - 2013", #LD
                       "2004 - 2019", #Ebro
                       "2002 - 2020", #Naiades
                       "2005",        #Cantabria
                       "2015 - 2017",   # Picos
                       "2005",        # Ecosurv
                       "1978 - 2004",   #RIVPACS
                       "2002 - 2003",   # STAR
                       "2000 - 2008",   # WISER
                       "2000",        # AQUEM SWEDEN
                       "2000 - 2013",   # Kaiser Lena Huttunen
                       "2007 - 2013",   # PUP
                       "1978 - 2017", # Edwin Peters
                       "2005 - 2018"    # Peter Haase
                       )
# 
impact$Area = c("Germany",     # LD
                "Spain",       # EBRO
                "France",      # NAIADES
                "Spain",       # Cantabria
                "Spain",       # Picos
                "Hungary",     # ECOSURV
                "UK",          # RIVPACS
                "Europe",      # STARS
                "Europe",      # WISER
                "Sweden",      # LEONARD SANDIN
                "Finnland",    # KLH
                "France",      # PUP
                "Netherlands", # EDWIN PETERS
                "Germany"      # PETER HAASE
                )
impact$`Number of Samples` = c(65211, # LD
                               3668,  # EBRO
                               27052, # NAIADES
                               55,    # Cantabria
                               24,    # Picos
                               491,   # ECOSURV
                               2504,  # RIVPACS
                               91,    # STARS
                               2565,  # WISER
                               150,   # LEONARD SANDIN
                               322,   # KLH
                               2694,  # PUP
                               44702,  # Edwin Peters
                               3044  # Peter Haase
                               )
impact$`Exclusion criterion` = c(
        "anthropogenic land use > 20\\%",
        paste0("IASPT", footnote_marker_symbol(1), "< 4.5 \\& Land use"),
        paste0("BDI",   footnote_marker_symbol(2), "< 14.5 \\& Land use"),
        "anthropogenic land use > 20\\%",
        "anthropogenic land use > 20\\%",
        "anthropogenic land use > 20\\%",
        #
        "original site selection",
        # RIVPACS
        "classification in data set \\& anthropogenic land use > 20",
        #STARS
        "anthropogenic land use > 20\\%",
        # WISER
        "anthropogenic land use > 20\\%",
        # LEONARD SANDIN
        "original site selection",
        # KLH
        "qualitative ratings of multiple anthropogenic stressors",
        # PUP
        "anthropogenic land use > 20\\%",
        # Edwin Peters
        "anthropogenic land use > 20\\%"  
        # Peter Haase
)
# 
impact$`Reference` = c(
                                       "Bhowmik \\& Schäfer (2015), Berger et al. (2018), Le et al. (2021)", # LD
                                       "Escribano et al (2018)",# EBRO
                                       "",# NAIADES
                                       "Alvarez et al. (2010), Alvarez et al. (2011)", # Cantabria
                                       "",# Picos
                                       "Schmera \\& Baur (2011), Schmera et al. (2012)",# ECOSURV
                                       "Clark et al. (2003), Turley et al. (2014)",# RIVPACS
                                       "Verdonschot (2006a), Johnson et al. (2007)",# STARS
                                       "Lyche-Solheim et al. (2013)",# WISER
                                       "Verdonschot (2006a)",# LEONARD
                                       "Huttunen et al. (2017)",# KLH
                                       "Mondy \\& Usseglio-Polaterra (2013), Alric et al. (2021)",# PUP
                                       "Peters et al. (2013)",# EDWIN
                                       "Haubrock et al. (2020)"# PETER
)

impact %>% 
        kbl(
                format = "latex",
                longtable = FALSE,
                caption = "Data sets used in the analysis. The provided number of samples contains unsuitable sites. The exclusion criterion gives the variables and thresholds we used to evaluate whether a site is impaired. For more detailed information on these criteria, see the supplementary information.",
                escape = F,
                booktabs = TRUE
        ) %>%
        kable_styling(
                latex_options = c("hold_position",
                                  "repeat_header"),
                full_width = TRUE,
                font_size = 7
        ) %>% 
        row_spec(0, bold = T) %>%
        footnote(symbol = c(
                "Iberian Average Score per Taxon; see Leunda et al. (2009) or Munné & Prat (2009)", 
                "Biological Diatom Index; see Lenoir & Coste (1996)") 
                ) %>% 
        column_spec(column = 1, width = "4cm") %>% 
        column_spec(column = 2, width = "1.5cm") %>% 
        column_spec(column = 3, width = "1.5cm") %>% 
        column_spec(column = 4, width = "1.5cm") %>% 
        column_spec(column = 5, width = "4cm") %>% 
        column_spec(column = 6, width = "4cm")  

```
\elandscape

\newpage

```{r fig-workflow, fig.cap="Scheme of the data analysis. Blue boxes represent data sets and yellow boxes analyses. Names of typologies have been abbreviated to BRT (broad river typology), GloRiC (Global River Classification), Illies (Illies freshwater ecoregions) and BGR (Biogeographic Regions)."}
readJPEG("fig/workflow.jpg") %>% 
  grid.raster()
```

The resulting data set had sufficient data to represent 13 of 20 broad river types.
We considered our samples representative for a river type, if we had more than 20 samples from the river type and the samples covered most of the area of the river type. 
<!-- Maps of the sampling locations for the individual river types are provided in the supplementary information. -->
We had insufficient samples of the river types: RT6, 7, 12, 13, 17, 19 and 20, consequently, these were omitted from further analyses.

```{r fig-map-used, fig.cap = "Map of samples used in the analyses. Each dot represents one sampling event. Darker dots indicate several overlapping sampling events which can be due spatially close sampling events or mutliple sampling events at one site (e.g. yearly samples)."}
readPNG("fig/map_sites_used.png") %>% 
  grid.raster()
```

<!-- How did we harmonize data sets.  -->
The data originated from different sources and required adjustments to ensure taxonomic consistency.
We used the Global Biodiversity Information Facility (GBIF) to replace synonyms with accepted names and to assign a taxonomic level to each observation.
To integrate the different data types, we transformed all data to presence-absence.
Next, we restricted observations to the classes Clitellata, Insecta, Malacostraca, Bivalvia and Gastropoda, as they occurred in all data sets. 
When harmonizing taxonomies across data sets, there is a trade-off between comparability among data sets and taxonomic resolution.
If we employ a single taxonomic level (e.g. genus) for all taxa across river types, we have to agglomerate observations with finer resolutions and omit those with coarser resolutions.
The resulting data set would be appropriate for comparisons between river types because taxa have consistent representations across river types.
If we vary the taxonomic resolution between river types and taxa, we artificially increase the dissimilarity between river types but attain the best available representation for the individual river types.
To compare freshwater typologies (see Comparison of typologies), we used the first approach and harmonized all observations at the genus level, omitting all observations with coarser taxonomic levels.
This resulted in a data set with `r data_genus_n_genus` genera.
To established typical assemblages (see Typical assemblages), we used the second approach. 
Typical assemblages were derived independently for each type, which allowed us to set an appropriate taxonomic level per river type and taxon.
For each taxon, we calculated the percentage of observations represented at each finer taxonomic resolution.
If this value was below 50 %, we harmonized all observations to this level.
If it was above 50%, we repeated the procedure with the next finer level.
We also included all observations with coarser than optimal resolution.
The resulting data set contained `r data_all_n_spe` taxa represented at the species level, `r data_all_n_gen` taxa represented at genus level and `r data_all_n_fol` taxa represented at family or coarser taxonomic levels.

## Comparison of typologies

### General approach 
@LycheSolheim2019 proposed two versions of the BRT typology.
BRT20 differentiates between 20 different river types, some of which are very rare.
To reduce the number of broad river types, @LycheSolheim2019 merged rare types with common types that matched the rare type in at least two type descriptors.
The resulting typology (BRT12) has twelve types.
Previous publications have used both (e.g. @Posthuma2020 used BRT12 and @Poikane2019 used BRT20) so we consider both.

To evaluate the two versions of BRT, we compared them with the three other freshwater typologies: i) the k-means typology of the global river classification framework [GloRiC, @OuelletDallaire2019]; ii) the freshwater ecoregions proposed by @illies1978limnofauna; and iii) the biogeographical regions proposed by the European Environmental Agency [BGR, @EuropeanEnvironmentalAgencyEEA2016].
These typologies can be categorized into: reach-based typologies (BRT and GloRiC) and regional typologies (Illies and BGR).
Reach-based typologies classify single river reaches.
The instances of individual types are not spatially contiguous and can be far apart.
They are commonly very close to instances of different types.
By contrast, regional typologies classify large contiguous areas and there is only one instance per type.
Different types are only adjacent at the boundaries between two types.
As with BRT before, we retained those types for which we had more than 20 samples.
See table \@ref(tab:tab-types) for the remaining types in each typology.

For each typology, we computed four cluster validity metrics, each emphasizing a different aspect of cluster quality: the average silhouette width, the Calinski-Harabasz index, the indicator value score and the classification strength.
We used the Jaccard distance between sites as a distance metric.

As a reference frame for these four metrics, we created two additional typologies.
To have an upper bound of what to expect, we clustered the sampling sites with an agglomerative clustering method.
This typology is constrained neither by environment nor by space but only by the biological data.
Thus, we expect it to delineate more clearly between different biological assemblages than the other typologies and to achieve better cluster validity scores.
We will refer to this typology as the biological typology. 
All agglomerative clustering methods begin with as many clusters as observations and then they iteratively combine the most similar observations into clusters until only one cluster remains.
The approaches differ in how this similarity is computed.
@lance1967general clarified that many approaches can be expressed as different parametrizations of one method: flexible beta clustering.
We computed biological typologies with different number of types and a $\beta$ parameters of 0.25. 
Based on two cluster validity metrics, we decided to use nine types. 
We refer to the supplementary information for more detailed information.
As a lower bound of what to expect, we created random typologies.
For each random typology, we first drew the number of types as a random variable from the interval between the lowest number of types in any of the typologies (6 in BGR) and the highest number (14 in GloRiC).
Then we assigned each sample randomly to one of the types.
We repeated this 100 times, to estimate how strongly cluster validity metrics would vary between random typologies. 
Each cluster validity criterion was computed for each of random typology. 
For each validity metric, we took the mean value of all 100 random typologies as a lower bound. 
The differences between the types of these typologies are purely random and hence 
there should be no systematic differences between them.
Therefore we expect low mean cluster validity scores relative to the other typologies.

### Cluster validity metrics  

Here, we provide a short primer on each cluster validity metric we used.
More extended information is provided in the supplementary information.
The average silhouette width [ASW, @Kaufmann1990] is the average of all silhouette widths.
A silhouette width is the difference between the average distance of sample $i$ to all samples from the most similar type and the average distance of sample $i$ to all other samples from the same type.
The silhouette width is scaled to lie between -1 and 1.
Values close to 1 indicate that most samples are more similar to samples from their own type than to those of other types.
Values around zero suggest that the sample lies between two types and values close to -1 signal that the sample is more similar to samples from another type than to those from its own.
In addition to the common ASW, we computed two modified versions, which summarize the distances in $a_i$ and $b_i$ with the generalized mean instead of the arithmetic mean [@Lengyel2019]. 
The generalized mean has one modifiable parameter, $p$, and for $p=1$ the modifiable mean is equal to the arithmetic mean.
As we increase $p$ the weight of outlying samples increases. 
Calculating the ASW for a range of $p$, allows us to evaluate cluster separation and compactness separately.  
We calculated the minimum distance ASW ($p=-\infty; ASW_{min}$) and the maximum distance ASW ($p= \infty; ASW_{max}$). 
 
<!-- What is Calinski-Harabasz? -->
The Calinski-Harabasz Index [CH, @calinski1974dendrite] is the scaled ratio of the distances between types and the distances within types.
Higher values indicate that distances between types are larger than those within types and thus that samples are more similar to samples from their own type than to samples from other types.  
<!-- What is the indicator value score? -->
The indicator value score (IVS) is based on the indicator value (IndVal) proposed by @dufrene1997species.
IVS is the fraction of taxa that are statistically significant indicators (at a significance level of 0.01) for some type of a typology.
High scores show that the commonness and specificity of taxa differ between types. 
Therefore higher scores indicate better typologies.  

<!-- What is Classification strength -->
Lastly, we computed the classification strength [CS, @VanSickle1997]. Classification strength is the difference between mean within cluster similarity ($\overline{W}$) and mean between cluster ($\overline{B}$) similarity.
We calculated similarities as 1 - Jaccard distance. 
The CS ranges between 0 ($\overline{W} = \overline{B}$) and 1 ($\overline{B} = 0$), where higher values indicate a stronger classification. 
CS has been used in similar analyses [e.g. @Gerritsen2000; @VanSickle2000; @Vasconcelos2013], which enables us to directly compare our results with those of previous studies.

We ranked the typologies based on each of these four cluster validity metrics,
from 1 to 5 relating to the worst and best performances respectively.
Differences smaller than 5% of the range between the biological and the mean of the random typologies were regarded as ties.
When typologies were tied, they all received the mean of their untied scores.
For example, if three typologies are tied for the first place, instead of 5, 4 and 3 points, each receives a score of 4.

## Typical assemblages

We established typical assemblages for the types of  all five freshwater typologies using the data set with mixed taxonomic levels. 
We define a typical taxon as one that commonly occurs in a type and typical assemblages as the lists of these taxa. 
Typical taxa differ from indicator taxa which usually need to be common and type-specific.
As a measure of commonness, we used the percentage of samples from a type that included a taxon.
We considered species as typical if they occurred in more than 20% of samples.
Coarser taxonomic levels needed to be more common to be typical.
Genera needed to occur in at least 33% of samples and families or coarser taxonomic levels in at least 66%.
Having derived typical assemblages that way, we assessed how similar they were within typologies.
Low similarities show that the common taxa differ between types, whereas high similarities imply that the same taxa frequently occur irrespective of type. 
As similarity metric, we used the Jaccard similarity.
When the similarity between type A and type B is 0.5, half of the taxa in type A's typical assemblage also occur in type B's.
Note that these similarities are asymmetric since they relate to the number of taxa in a typical assemblage. 
If type B's typical assemblage has more taxa, the similarity between B and A will be lower than that between A and B. 
<!-- Why do we do this ? -->  

Lastly, we derived non-redundant typical assemblages. 
Two typical assemblages were considered redundant when their similarities exceeded 0.7 both ways.
We iteratively combined the most similar river types until no redundant types remained.
For example, the BRT20-types RT8 and 9 (mid-altitude siliceous rivers) were redundant and combined into RT8_9.
Subsequently, RT8_9 was found to be redundant with RT16 (glacial rivers) so these types were combined to RT8_9_16. 
This way, we can evaluate whether the combinations of types from BRT20 to BRT12 are justified by biological homogeneity.
In the supplementary information, we provide lists of all complete and non-redundant typical assemblages. 

<!-- Software  -->

## Data and Software

As geospatial representation of the BRT typology, we used the data provided by @globevnik2019btfr (Accessed 11.12.2019). 
The version of the GloRiC typology we used is available under https://www.hydrosheds.org/page/gloric (Accessed 30.07.2019). 
Illies freshwater ecoregions were queried from https://www.eea.europa.eu/data-and-maps/data/ecoregions-for-rivers-and-lakes  and the EEA biogeographical regions from https://www.eea.europa.eu/data-and-maps/data/biogeographical-regions-europe-3 (both Accessed 08.01.2020). 
We used the Catchment Characterization and Modeling database [@Vogt2007] and the Corine Landcover data from 2018 [@clc18] to derive the relative area of land use types in catchments.

All computations were conducted in the R Statistical Environment v. 4.0.3 [@RCT2020]. Data were prepared using data.table 1.14.0 [@Dowle2021], tidyverse packages [@Wickham2019] and taxize 0.9.98 [@ScottChamberlain2013; @Chamberlain2020]. Geospatial analyses were conducted using sf [@Pebesma2018]. 
Clusters were created and evaluated with fpc [@Hennig2020], indicspecies [@Caceres2009]. 
Generalized silhouette widths were computed with the R functions provided in the supplementary information of @Lengyel2019.
Figures and maps were created with ggplot2 [@Wickham2016] and tmap [@Tennekes2018].
