\newpage
\blandscape
# Supplementary Materials 
```{r }
impact = matrix(nrow = 14, ncol = 6)
impact %<>% as.data.frame()
names(impact) = c("data set", "Area", "Time Span", "Number of Samples", "criterion", "connected publications")
impact$`data set` = c("Monitoring data from German federal agencies", 
                      "Ebro Hydrographic Confederation", 
                      "Naiades",
                      "Cantabria",
                      "Picos de Europa",
                      "Ecosurv",
                      "RIVPACS database",
                      "STAR",
                      "WISER",
                      "AQUEM Sweden",
                      "Koutajoki drainage basin",
                      "Monitoring data from the RCS national network managed by the French water agencies",
                      "Monitoring data from Dutch regional water authorities ",
                      "Monitoring data from the German Federal state of Hesse")

impact$`Time Span` = c("1968 - 2013", #LD
                       "2004 - 2019", #Ebro
                       "2002 - 2020", #Naiades
                       "2005",        #Cantabria
                       "2015-2017",   # Picos
                       "2005",        # Ecosurv
                       "1978-2004",   #RIVPACS
                       "2002-2003",   # STAR
                       "2000-2008",   # WISER
                       "2000",        # AQUEM SWEDEN
                       "2000-2013",   # Kaiser Lena Huttunen
                       "2007-2013",   # PUP
                       "1978 - 2017", # Edwin Peters 
                       "2005-2018"    # Peter Haase 
                       )

impact$Area = c("Germany",     # LD
                "Spain",       # EBRO
                "France",      # NAIADES
                "Spain",       # Cantabria
                "Spain",       # Picos
                "Hungary",     # ECOSURV
                "UK",          # RIVPACS
                "Europe",      # STARS
                "Europe",      # WISER
                "Sweden",      # LEONARD SANDIN
                "Finnland",    # KLH
                "France",      # PUP 
                "Netherlands", # EDWIN PETERS 
                "Germany"      # PETER HAASE 
                )
impact$`Number of Samples` = c(65211, # LD
                               3668,  # EBRO
                               27052, # NAIADES
                               55,    # Cantabria
                               24,    # Picos
                               491,   # ECOSURV
                               2504,  # RIVPACS
                               91,    # STARS
                               2565,  # WISER
                               150,   # LEONARD SANDIN
                               322,   # KLH
                               2694,  # PUP
                               44702,  # Edwin Peters 
                               3044  # Peter Haase 
                               )
# impact$availability = c(
#         "not openly available",
#         "Link 1",
#         "Link 2",
#         "not openly available",
#         "not openly available",
#         "not openly available",
#         "Link 3",
#         "not openly available",# STARS
#         "not openly available",# WISER
#         "not openly available",# Leonard Sandin
#         "not openly available",# KLH
#         "not openly available",# PUP
#         "Link 4",# Edwin Peters
#         "not openly available" # Peter Haase
# )
impact$criterion = c("Land use", 
                     "IASPT > 4.5 & Land use",
                     "BDI   > 14.5 & Land use",
                     "Land use",
                     "Land use",
                     "Land use", # 
                     "original site selection", # RIVPACS
                     "classification in data set & Land use", #STARS
                     "Land use", # WISER
                     "Land use", # LEONARD SANDIN
                     "original site selection", # KLH
                     "qualitative ratings of multiple anthropogenic stressors", # PUP
                     "Land use", # Edwin Peters 
                     "Land use"  # Peter Haase
)

impact$`connected publications` = c(
                                       "Bhowmik & Schäfer (2015), Berger et al. (2018), Le et al. (2021)", # LD
                                       "Escribano et al (2018)",# EBRO
                                       "",# NAIADES
                                       "Alvarez et al. (2010), Alvarez et al. (2011)", # Cantabria
                                       "",# Picos
                                       "Schmera & Baur (2011), Schmera et al. (2012)",# ECOSURV
                                       "Clark et al. (2003), Turley et al. (2014)",# RIVPACS
                                       "Verdonschot (2006a), Johnson et al. (2007)",# STARS
                                       "Lyche-Solheim et al. (2013)",# WISER
                                       "Verdonschot (2006a)",# LEONARD
                                       "Huttunen et al. (2017)",# KLH
                                       "Mondy & Usseglio-Polaterra (2013), Alric et al. (2021)",# PUP
                                       "Peters et al. (2013)",# EDWIN
                                       "Haubrock et al. (2020)"# PETER
)
# impact_hux =
#         impact %>%
#         dplyr::select(`data set`, `Time Span`) %>%
#         as_hux()
# escape_contents(impact_hux) = FALSE
# impact_hux
# impact_hux =
#         impact %>%
#         as_hux(add_colnames = TRUE) %>%
#         set_wrap(TRUE) %>%
#         set_bold(1, everywhere) %>% 
#         set_bottom_border(1, everywhere) %>% 
#         set_width(1.5) %>% 
#         set_number_format(0) %>%
#         set_font_size(6) %>% 
#         #set_right_border(everywhere, everywhere, brdr(1, "solid", "black")) %>% 
#         set_col_width(col = 1, value = .2) %>% 
#         set_col_width(col = 2, value = .07) %>% 
#         set_col_width(col = 3, value = .08) %>% 
#         set_col_width(col = 4, value = .08) %>% 
#         set_col_width(col = 5, value = .2) %>% 
#         set_col_width(col = 6, value = .2) 
# 
# caption(impact_hux) = "Overview of data sets used for the analysis. "
# # The respective data sets are available under link 1: www.datossuperficiales.chebro.es:81/WCASF/?rvn=1 ; link 2: www.naiades.eaufrance.fr/france-entiere#/ ; link 3: www.ceh.ac.uk/services/rivpacs-reference-database , link 4: waterkwaliteitsportaal.nl"
# impact_hux
```

```{r tab-impact-criteria}
impact %>% 
        kbl(
                format = "latex",
                longtable = FALSE,
                caption = "test ",
                booktabs = TRUE
        ) %>%
        kable_styling(
                latex_options = c("hold_position",
                                  "repeat_header"),
                full_width = TRUE,
                font_size = 7
        ) %>% 
        row_spec(0, bold = T) %>% 
        column_spec(column = 1, width = "4cm")
```

\elandscape

\newpage

In table \@ref(tab:tab-impact-criteria), *Land use* refers to the 20% cutoff for relative area of agriculture or urban areas in a catchment. This criterion was applied wherever no other information was available. 
IASPT is short for Iberian Average score per taxon which refers to the average Iberian biological monitoring working party score (IBMWP, between 1 and 10) of a sample. 
The IASPT has been shown to be a reliable indicator of water quality [e.g. @Leunda2009; @Munne2009]. 
The Biological Diatom Index (BDI) [@lenoir1996development] is a standardized way to evaluate the quality of freshwater systems in France using Diatoms.  
In two of the data sets (RIVPACS and Koutajoki), the sampling sites were selected to represent reference conditions so we used all of the sites. 
For the data from the RCS national network several qualitative metrics were available. 
The measured stressors were: organic matter, nitrogen compounds (except nitrates), nitrates, phosphorous compounds, suspended matter, organic micropollutants (other), mineral micropolluants (metal), pesticides and polycyclic aromatic hydrocarbons (PAH). 
Each was rated on a scale of high, good, intermediate, poor and bad. 
All sites for which more two or more categories were rated as intermediate or one more were rated as either poor or bad was removed from the analysis. 

\newpage

```{r tab-types}
data = readRDS("data/12_sxs_genus_typology_wo_bio.rds")
#data = readRDS("../../data/12_sxs_genus_typology_wo_bio.rds")
gloric_unique = data %>% pull("gloric") %>% unique %>% sort
brt12_unique  = data %>% pull("ls12") %>% unique %>% sort
brt12_unique = brt12_unique[c(1,4:11,2,3)]
brt20_unique  = data %>% pull("ls20") %>% unique %>% sort
brt20_unique = brt20_unique[c(1,9:13,2:8)]
illies_unique = data %>% pull("illies") %>% unique %>% sort
eea_unique    = data %>% pull("eea") %>% unique %>% sort

s1 = data.table(typology = c("BRT12", "BRT20", "GloRiC", "Illies", "BGR"), 
           types    = c(
                   paste(brt12_unique, collapse = ",\ "), 
                   paste(brt20_unique , collapse = ",\ "),     
                   paste(gloric_unique, collapse = ",\ "),     
                   paste(illies_unique, collapse = ",\ "),     
                   paste(eea_unique, collapse = ",\ ")     
                        )
           )
s1_hux =
        s1 %>%
        as_hux(add_colnames = TRUE) %>%
        set_wrap(TRUE) %>%
        set_bold(1, everywhere) %>% 
        set_bottom_border(1, everywhere) %>% 
        set_width(1) %>% 
        set_number_format(0) %>%
        set_font_size(7) 
        #set_right_border(everywhere, everywhere, brdr(1, "solid", "black")) %>% 
        # set_col_width(col = 1, value = .2) %>% 
        # set_col_width(col = 2, value = .07) %>% 
        # set_col_width(col = 3, value = .08) %>% 
        # set_col_width(col = 4, value = .08) %>% 
        # set_col_width(col = 5, value = .2) %>% 
        # set_col_width(col = 6, value = .2) 
caption(s1_hux) = "Types of different typologies that were used in the analyses."
s1_hux
```

\newpage

```{r tab-sm-typicals}
ta_table = readRDS("data/18_ta_table_for_paper.rds")
#ta_table = readRDS("../../data/18_ta_table_for_paper.rds")
# sig_id =map(.x = 2:ncol(ta_table),
#     .f = ~ str_detect(pull(ta_table[,.x]), "\\*")
# )

ta_table %>%
        kbl("latex",
            longtable = TRUE,
            caption = "Typical assemblages. Values give the proportion of sites of broad tiver type where the taxon is present. Combinations for which a taxon is found to be typical are marked with a (*).",
            booktabs = TRUE
            ) %>%
        kable_styling(
                latex_options = c("hold_position",
                                  "repeat_header"),
                full_width = TRUE,
                font_size = 9
        ) %>%
        row_spec(0, bold = T) 
```

\newpage
```{r tab-sm-indi}
indi = readRDS("data/15_indicator_list_w_non_significant.rds")
#indi = readRDS("../../data/15_indicator_list_w_non_significant.rds")
names(indi) = paste0("RT", c(1,2,3,4,5,8,9, 10, 11, 14, 15, 16, 18))
for (i in 1:length(indi)){
                indi[[i]]$rt = names(indi)[i]
}
## - drop empty types 
indi2 <- rbindlist(indi)

indi2 %>%
        arrange(taxon) %>%
        filter(taxon != "Hydracarina") %>% 
        mutate(indval = round(indval, 2))  %>% 
        mutate(indval_chr = as.character(indval))  %>% 
        mutate(indval_chr = case_when(indicator ~ paste0(indval_chr, "*"),
                                      !indicator ~ indval_chr))  %>% 
        pivot_wider(id_cols = taxon,
                    names_from = rt,
                    values_from = indval_chr) %>%
        relocate(RT3, .before = RT4) %>%
        relocate(RT2, .before = RT3) %>%
        relocate(RT1, .before = RT2) %>%
        relocate(RT8, .before = RT9) %>%
        relocate(RT10, .before = RT11) %>% 
        mutate(across(!taxon, function(x)ifelse(is.na(x),0,x))) ->
        indi3

## -- which rows contain taxa that are indicative for any stream type 
indi3$id = apply(indi3, 1, function(x) any(str_detect(string = x, pattern = "\\*")))

## -- make kable table 
indi3 %>%
        ## -- only genera that are indicators
        filter(id == TRUE) %>%
        ## -- drop id column 
        dplyr::select(!id) %>% 
        ## -- kbl as wrapper for kable()
        kbl(
                format = "latex",
                longtable = TRUE,
                caption = "Indicator scores for all taxon and stream type combination where at least one is statistically                            significant. Statistically significat entries are marked with a *. ",
                booktabs = TRUE
        ) %>%
        kable_styling(
                latex_options = c("hold_position",
                                  "repeat_header"),
                full_width = TRUE,
                font_size = 7
        ) %>% 
        row_spec(0, bold = T) %>% 
        column_spec(column = 1, width = "4cm")
```

```{r old hux code typical}
#ta_table = readRDS("../../data/18_ta_table_for_paper.rds")
# hux_ta_table = as_huxtable(ta_table)
# hux_ta_table %<>%
#         map_background_color(
#                              fn = by_ranges(breaks = c(0.2),
#                                             values = c("white", "green"))) %>%
#         # map_text_color(
#         #                fn = by_values("1" = "green", "0" = "white")) %>%
#         set_bold(1, everywhere) %>% 
#          set_font_size(7) %>% 
#         set_all_borders(1) %>% 
#         set_row_height(.0001)
# hux_l = split_across(hux_ta_table, after = 25)
# hux_l[[1]]
# hux_l[[2]]
```

